---
title: "Permutation Testing"
author: "Kevin Ferris"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Permutation Testing}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = T, 
                      message = F, 
                      warning = F, 
                      tidy = F, 
                      comment = "#>", 
                      eval = F)
```


This project uses permutation testing to compare density surfaces.  This testing was designed to work with the PITCHf/x data, but it could be easily customized to work with any dataset.  The workhorse function is `perm_pvals`.  This calculates p-values comparing for the observed difference in density surfaces at each of the $n \times n$ grid points.

##  A Simple Example

This compares the density surface for responsible pitchers to the density surface for pitchers who are not responsible.  

```{r example, eval=TRUE}
library(Kwproj)
data(pit_LvL)
simple_pvals <- perm_pvals(15, 10, pit_LvL)
```

## The Process

1) calculate the observed difference in densities
```{r obs-diff}
npoints <- 15 # number of grid points to calculate densities
B <- 10       # number of permutations
obs_diff <- dens_diff(npoints, data = pit_LvL)
```

2) Run the permutations and calculate the p-values.  This is now one step.  `perm_pvals` calls on `perm_dens` to perform the permutations and `calc_pvals` to compare the permutations to the observed data.
```{r perm-pvals}
pvals <- perm_pvals(B, npoints, data = pit_LvL, obs_diff = obs_diff)
```

3) Here is where we'll want to make sure to save the object.  
```{r save-pvals, eval=FALSE}
save(pvals, file = "my-location.RData")
```

4) If you want, you can do some fun visualizations.  Here we plot the observed p-values by location

```{r plots, message = FALSE}
library(ggplot2)
library(scales)
pvals2 <- pvals_to_long(pvals) 

# actual size
qplot(row, col, data = pvals2, colour = pval) + 
  scale_colour_gradient2(low = muted("red"), mid = "grey", high = muted("blue"), 
                         midpoint = .2)

# whether they're greater than 0.05
pvals2 <- pvals2 %>% 
  mutate(size = ifelse(pvals < .05, "mouse", "elephant") %>% 
           factor(levels = c("mouse", "elephant")))
qplot(row, col, data = pvals2, colour = size) + 
  labs(x = "Horizontal Location", y = "Vertical Location") + 
  theme_bw()
```


## Vignette Info

Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
