---
title: "Permutation Testing"
author: "Kevin Ferris"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Permutation Testing}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = T, 
                      message = F, 
                      warning = F, 
                      tidy = F, 
                      comment = "#>", 
                      eval = F)
```


This project uses permutation testing to compare density surfaces.  This testing was designed to work with the PITCHf/x data, but it could be easily customized to work with any dataset.  The workhorse function is `perm_pvals`.  This calculates p-values comparing for the observed difference in density surfaces at each of the $n \times n$ grid points.

##  A Simple Example

This compares the density surface for responsible pitchers to the density surface for pitchers who are not responsible.  Here, we perform 15 permutations using 10 grid points.

```{r example, eval=TRUE}
library(Kwproj)
data(pit_LvL)
simple_pvals <- perm_pvals(15, 10, pit_LvL)
simple_pvals
```

## The Process

1) calculate the observed difference in densities
```{r obs-diff}
npoints <- 15 # number of grid points to calculate densities
B <- 10       # number of permutations
obs_diff <- dens_diff(npoints, data = pit_LvL)
```

2) Run the permutations and calculate the p-values.  This is now one step.  `perm_pvals` calls on `perm_dens` to perform the permutations and `calc_pvals` to compare the permutations to the observed data.
```{r perm-pvals}
pvals <- perm_pvals(B, npoints, data = pit_LvL, obs_diff = obs_diff)
```

3) If you want, you can do some fun visualizations.  Here we plot the observed p-values by location

```{r plots, message = FALSE}
library(ggplot2)
library(scales)
pvals2 <- pvals_to_long(pvals) 

# actual size
qplot(row, col, data = pvals2, colour = pval) + 
  scale_colour_gradient2(low = muted("red"), mid = "grey", high = muted("blue"), 
                         midpoint = .2)

# whether they're greater than 0.05
pvals2 <- pvals2 %>% 
  mutate(size = ifelse(pvals < .05, "mouse", "elephant") %>% 
           factor(levels = c("mouse", "elephant")))
qplot(row, col, data = pvals2, colour = size) + 
  labs(x = "Horizontal Location", y = "Vertical Location") + 
  theme_bw()
```

